---
title: "Empirical analysis on Exponential distribution and Central Limit Theorem"
author: "Alessandro Rotatori"
date: "27 giugno 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis

The objective of the following empirical analysis is to describe the characteristics of the exponential distribution with respect to the simple statistics, mean and variance. Furthermore, the validity of Central Limit Theorem for estimating the population mean is evaluated.

## Analysis

The Empirical Analysis is performed by analysing the behavior of the exponential distribution when simulating 40 iid exponentials.

The parameters relevant for the analysis, therefore, are the following:

- *lambda* = 0.2, parameter of the exponential distribution;
- *n*, number of iid exponential variables;
- *sim*, number of simulations.


```{r setting_param}
lambda <- 0.2
n <- 40
sim <- 1000
mn <- 1 / lambda
sd <- 1 / lambda
```

The empirical test, as seen

## Appendix
